---
title: EDA
output: github_document
---

This is a exploratory data analysis to explore the [Marshall Violent Crime](https://github.com/themarshallproject/city-crime) dataset.

```{r}
# load libraries
suppressPackageStartupMessages(library(tidyverse))
```

```{r}
# load data
crime <- read_csv("../data/ucr_crime_1975_2015.csv")
```

```{r}
# size of data
dim(crime )
```

There are 2,829 entries in the raw dataset.

```{r}
# explore department_name (city name) variable
crime$department_name %>%
  unique()
```

There are some potential issues with this variable:

- some police department oversees counties as well, for the purpose of this app, we only want to display information cities.

- the variable value "National" is not a meaningful entry for our app.

```{r}
# count the number of problematic city names
crime$department_name %>%
  unique() %>% 
  str_subset(pattern = "County|National")
```

There are 9 potentially problematic city names for our application.

```{r}
# explore incomplete entries
# identify annual reports with less than 12 months reported
crime %>% 
  filter(months_reported < 12)
```

There are 53 annual reports with less than 12 months reported.
